Integer Programming from Quantum Annealing and Open Quantum Systems

### 2020-09-23 paper/main.tex
Updated Linblad eq indices and final discussion

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
 \partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \mathcal{L}(\rho(t), H(t))
\end{align}
where $\rho (t)$ is the density matrix at time $t$, $H(t)$ is the time-dependent Hamiltonian and $\mathcal{L}$ is the Linblad operator implementing the decoherence models.
In this work, we consider two types of decoherence models: full-counting statistics~\cite{PhysRevE.90.022103,RevModPhys.81.1665} and single-qubit amplitude damping (local damping)~\cite{10.5555/1972505,preskill1998lecture}.
The full-counting statistics term models the global decoherence to all the qubits due to the classical reservoir.
The local damping term models the decoherence of each qubit independently.
Details of the master equation are given in Sec.~\ref{sec:methods:lindblad}.
Implementation of the annealing schedule and annealing offsets for the simulation are discussed in Sec.~\ref{sec:methods:annealing-schedule}.

@@

The minimal energy solution to Eq.~\eqref{eq:initial-ip-def} and Eq.~\eqref{eq:ilp-slack-bit-energy} are exactly the same if the penalty term is larger than the energy gap of the first excited solution: $p > E_1 - E_0$.
Thus some knowledge of the problem is required.
In principle, it is possible to set the penalty term arbitrarily large, at the cost of problem resolution: large values of For $p$ increase the highest available energy of the system by multiples of $p$.
After normalization of the QUBO, this corresponds to decreasing the energy gap between the ground state and the first excited state.
Thus, if solvers have finite precision, one must estimate reasonable values for $p$: for large $p$ more solutions fulfill the constraints, while for small $p$, more solutions minimize the objective function.

@@

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we evaluate the master equation in Lindblad form
\begin{align}
 \partial_t \rho (t) = &  \frac{-i}{\hbar} \left[H(t) , \rho(t)\right] \nonumber \\
 & + \sum_j \left(2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \}\right) ,
\end{align}
where $\rho (t)$ is the density matrix at time $t$.
$H(t)$ is the time-dependent Hamiltonian
\begin{align}
 \label{eq:annealH}
 H_{anneal}(t)  = & - \sum_i  A_i(t)\sigma^x_i +\sum_i B_{i}(t) h_i \sigma^z_i \notag \\
 & + \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j  ,
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions.
The site dependency takes into account of the annealing offset.
$[,]$ denotes the Lie bracket.
$L_j$ are the Lindblad operators for decoherence.
$\{, \}$ denotes the anti-commutator.

For full-counting statistics, the Lindblad operator is
\begin{align}
\mathcal{L}[\rho (t)] = & \Gamma_{\textrm{fc}} \sum_j   \left(2 S_j \rho(t) S_j^\dagger - \{ S^\dagger_j S_j, \rho(t) \}\right) \notag \\
& + e^{-\beta \Delta E_j} \left(2 S_j^\dagger \rho(t) S_j - \{ S_j S^\dagger_j, \rho(t) \} \right)
\end{align}
where $j$ is the index for the inter-level spacing $\Delta E_j=E_\mu-E_\nu>0$. $S_j=|E_\nu \rangle \langle E_\mu|$ denotes the many-body lowering operator.
$\{, \}$ denotes the anti-commutator. $\Gamma = 1/T_c$ is the decoherence rate for coherence time $T_c$.
That is, due to the interaction with the classical thermal bath, there is a probability that the system hops from each higher-energy many-body state to a lower-energy many-body state.
The probability of the inverse process is given by a Boltzmann factor.
To model the local decoherence of each qubits in the non-interacting limit, we also consider the amplitude damping for non-interacting qubits.
The Lindblad operator is
\begin{align}
\mathcal{L}[\rho (t)] = &  \Gamma_{\textrm{loc}} \sum_j \left(2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \} \right)\notag \\
& + e^{- 2 \beta |h_j| }  \left(2L_j^\dagger \rho(t) L_j - \{ L_j L_j^\dagger, \rho(t) \} \right) ,
\end{align}
where $j$ is the index for qubit.
$L_j= \sigma^{+}_j=|1\rangle \langle 0|$ for $h_j>0$ and $L_j= \sigma^{-}_j=|0\rangle \langle 1|$ for $h_j<0$.
That is, each qubit is damped toward its local ground state if we ignore all the interactions.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Annealing schedule and offsets}
\label{sec:methods:annealing-schedule}
%----------------------------------------------------------------------------------------
In this section, we discuss the details of the annealing schedule with respect to the dimensionless normalized time $s$.
On D-Wave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits ($B(s) \to B_i(s)$).
In Fig.~\ref{fig:anneal_schedule}, the default D-Wave annealing schedule is shown in black, in addition to the effects of applying negative offsets (effective time delay) to $A(s)$ and $B(s)$ in blue.
Further documentation is provided by D-Wave in Ref.~\cite{dwave_as, dwave_as_docu}.

@@

The coefficients are nonlinear in time because the underlying control variable $c(s)$ which is designed to grow the persistent current $I_p(s)$ linearly in time.
The effective time delay is implemented by introducing an offset as $c(s) + \delta$.
As a result, systematic errors are introduced because the final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets.
Additionally, the values of the coefficients are unknown outside of $s\in [0, 1]$.
We extrapolate their values by a linear extrapolation, and only employ negative offsets such that this effect only enters at the beginning of the annealing process rather than the end.

We verify that our implementation of annealing offsets on the simulator is consistent with D-Wave by solving the following three qubit Hamiltonian
\begin{align}
	\label{eq:offset_test_hamiltonian}
	H^{\textrm{problem}} =
	\begin{pmatrix}
		-0.25 & 1 & 0 \\
		0 & -0.25 & 0 \\
		0 & 0 & -0.25
	\end{pmatrix}
\end{align}
which has a doubly-degenerate ground state of $(0, 1, 1)$ and $(1, 0, 1)$. An annealing offset is then applied to either qubit 0 or 1, and breaks the ground state degeneracy of the system. Because of the systematic error introduced when assigning an offset to a qubit, the final Hamiltonian given by Eq.~(\ref{eq:annealH}) will have small deviations from the input problem Hamiltonian. For example, we expect a negative offset to qubit 0 to Eq.~(\ref{eq:offset_test_hamiltonian}) will yield $(1, 0, 1)$ as the unique ground state given Eq.~(\ref{eq:annealH}).

>>>>

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
 \label{eq:sim:linblad-eq}
 \partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \mathcal{L}(\rho(t), H(t))
\end{align}
where $\rho (t)$ is the density matrix at time $t$, $H(t)$ is the time-dependent Hamiltonian and $\mathcal{L}$ is the Linblad operator implementing the decoherence models.
In this work, we consider two types of decoherence models: full-counting statistics~\cite{PhysRevE.90.022103,RevModPhys.81.1665} and single-qubit amplitude damping (local damping)~\cite{10.5555/1972505,preskill1998lecture}.
The full-counting statistics term models the global decoherence to all the qubits due to the classical reservoir.
The local damping term models the decoherence of each qubit independently.
Details of the master equation are given in Sec.~\ref{sec:methods:lindblad}.
Implementation of the annealing schedule and annealing offsets for the simulation are discussed in Sec.~\ref{sec:methods:annealing-schedule}.

@@

The minimal energy solution to Eq.~\eqref{eq:initial-ip-def} and Eq.~\eqref{eq:ilp-slack-bit-energy} are exactly the same if the penalty term is greater or equal to the energy gap of the first excited solution: $p \geq E_1 - E_0$.
Thus some knowledge of the problem is required.
In principle, it is possible to set the penalty term arbitrarily large, at the cost of problem resolution: large values for $p$ increase the highest available energy of the system by multiples of $p$.
After normalization of the QUBO, this corresponds to decreasing the energy gap between the ground state and the first excited state.
Thus, if solvers have finite precision, one must estimate reasonable values for $p$: for large $p$ more solutions fulfill the constraints, while for small $p$, more solutions minimize the objective function.

@@

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we evaluate the master equation Eq.~\eqref{eq:sim:linblad-eq} in Lindblad form.
The explicit time-dependence of the Hamiltonian is given by
\begin{align}
 \label{eq:annealH}
 H(t)  = & - \sum_i  A_i(t)\sigma^x_i +\sum_i B_{i}(t) h_i \sigma^z_i \notag \\
 & + \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j  ,
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions.
The site dependency takes into account of the annealing offsets.
The simulation takes two decoherance models into account.
Full-counting statistics and local decoherence.

For full-counting statistics, the Lindblad operator is
\begin{align}
\mathcal{L}_{\textrm{fc}}(\rho (t), H(t)) = & \Gamma_{\textrm{fc}} \sum_{j, i<j} \left[  \left(2 S_{ij} \rho(t) S_{ij}^\dagger - \{ S^\dagger_{ij} S_{ij}, \rho(t) \}\right) \notag \right.
\\
& \left. + e^{-\beta \Delta E_{ij}} \left(2 S_{ij}^\dagger \rho(t) S_{ij} - \{ S_{ij} S^\dagger_{ij}, \rho(t) \} \right)\right]
\end{align}
where $\{, \}$ denotes the anti-commutator, $(ij)$ is the index for the inter-level spacing $\Delta E_{ij}=E_j-E_i>0$,  $S_{ij}=|E_i \rangle \langle E_j|$ denotes the many-body lowering operator and $\Gamma_{fc} = 1/T_{fc}$ is the full-counting decoherence rate for coherence time $T_{fc}$.
That is, due to the interaction with the classical thermal bath, there is a probability that the system hops from each higher-energy many-body state to a lower-energy many-body state.
The probability of the inverse process is given by a Boltzmann factor.

To model the local decoherence of each qubits in the non-interacting limit, we also consider the amplitude damping for non-interacting qubits.
For the local decoherance model, the Lindblad operator is
\begin{align}
\mathcal{L}_{\textrm{loc}}(\rho (t), H(t)) = &  \Gamma_{\textrm{loc}} \sum_j \left(2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \} \right)\notag \\
& + e^{- 2 \beta |h_j| }  \left(2L_j^\dagger \rho(t) L_j - \{ L_j L_j^\dagger, \rho(t) \} \right) ,
\end{align}
where $j$ is the index for qubit.
$L_j= \sigma^{+}_j=|1\rangle \langle 0|$ for $h_j>0$ and $L_j= \sigma^{-}_j=|0\rangle \langle 1|$ for $h_j<0$.
That is, each qubit is damped toward its local ground state if we ignore all the interactions.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Annealing schedule and offsets}
\label{sec:methods:annealing-schedule}
%----------------------------------------------------------------------------------------
In this section, we discuss the details of the annealing schedule with respect to the dimensionless normalized time $s$.
On D-Wave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits (see E.~\eqref{eq:annealH}).
In Fig.~\ref{fig:anneal_schedule}, the default D-Wave annealing schedule is shown in black, in addition to the effects of applying negative offsets (effective time delay) to $A(s)$ and $B(s)$ in blue.
Further documentation is provided by D-Wave in Ref.~\cite{dwave_as, dwave_as_docu}.

@@

The coefficients $A(s)$ and $B(s)$ follow the underlying control variable $c(s)$, which is designed to grow the persistent current $I_p(s)$ linearly in time.
The effective time delay is implemented by introducing an offset as $c(s) + \delta$.
Because annealing schedules are finite, systematic errors are introduced because the final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets.
Additionally, the values of the coefficients are unknown outside of $s\in [0, 1]$.
We only employ negative offsets such that this unkonwn coefficient range, approximated by a linear extrapolation, only enters at the beginning of the annealing process.

We verify that our implementation of annealing offsets on the simulator is consistent with D-Wave by solving the following three qubit Hamiltonian
\begin{align}
	\label{eq:offset_test_hamiltonian}
	H^{\textrm{problem}} =
	\begin{pmatrix}
		-0.25 & 1 & 0 \\
		0 & -0.25 & 0 \\
		0 & 0 & -0.25
	\end{pmatrix}
\end{align}
which has a doubly-degenerate ground state of $(0, 1, 1)$ and $(1, 0, 1)$. An annealing offset is then applied to either qubit 0 or 1, and thus breaks the ground state degeneracy of the system. Because of the systematic error introduced when assigning an offset to a qubit, the final Hamiltonian given by Eq.~(\ref{eq:annealH}) will have small deviations from the input problem Hamiltonian. For example, we expect a negative offset to qubit 0 to Eq.~(\ref{eq:offset_test_hamiltonian}) will yield $(1, 0, 1)$ as the unique ground state given Eq.~(\ref{eq:annealH}).
